### 2018-12-08 writing/thesis/chapters/implementation.tex
Some more work.

\chapter{Scenarios}\label{sec:scenarios}
\section{Numerical Convergence Test}
We first define the $L_p$ norms for $p > 0$ by
% notation: https://en.wikipedia.org/wiki/Lp_space#Lp_spaces
\begin{equation}
  \label{eq:Lp-nrom}
  \Vert f(x) \Vert = \left( \int_K \vert f(x) \vert^p d\mu  \right)^{1/p}.
\end{equation}
We start by a known analytical solution $f(x)$ and compare it to our approximation $\hat{f}(x)$.
Let $Q$ and $\hat{Q}$ denote the analytical and approximate solution at node.
We start by computing the point-wise error, and compute the cell-wise error from this.
Observe that for our broken space $\Omega$
\begin{equation}
  \label{eq:lp-norm-broken}
 \Vert f(x) \Vert = \sum_{K \in \Omega} \Vert f_K(x) \Vert_p.
\end{equation}

We are now ready to integrate the error for each cell.
The output contains the position and conserved variables of each point.
Each cell consists of $(N + 1)^d$ nodes, each associated with a quadrature weight $w_i$.
The quadrature weights are associated with the reference cuboid.
We thus need to map our element to the reference element, before we can compute the integral.

Let $V$ denote the volume (or area) of each cell.
For example, in three dimensions $V = \Delta x \, \Delta y \, \Delta z$.
The final result is then
\begin{equation}
  \Vert f_K(x) \Vert = \left( V \sum_{\bm{i}} \vert f(\bm{x}_{\bm{i}}) \vert^p w_{{\bm{i}}}  \right)^{1/p},
\end{equation}
where we used the mapping to the reference triangle (integration by substitution).

Following~\cite{dumbser2010arbitrary}, the scenario can be described in primitive variables by setting
\begin{align}
  \pressure &= \pressure_0 \cos( \bm{k} \bm{x} - \omega t ) + \pressure_b, \\
  \Qrho &= \Qrho_0 \sin (\bm{k} \bm{x} - \omega t) + \Qrho_b, \\
  \Qv &= \bm{v_0} \sin(\bm{k} \bm{x} - \omega t).
\end{align}
We set the constants to \( \left(  \pressure_0 = 0.1, \pressure_b = \gamma^{-1}, \Qrho_0 = 0.5, \Qrho_b = 1, 
\bm{v_0} = 0.25 {(1,1)}^\intercal, \bm{k} = \pi/5 {(1,1)}^\intercal \right) \).

We derive a source term by inserting this solution into the \textsc{pde} by using the symbolic math toolkit \textit{Sage}~\cite{sagemath}.
This procedure is called method of manufactured solutions~\cite{salari2000code}.

\section{Classical Scenarios}

\subsection{The first Stokes Problem}
\cite{dumbser2010arbitrary}
\subsection{Taylor-Green Vortex}
\begin{align}
  \begin{split}
  \Qv(\bm{x}, t) &= \exp(-2 \mu t)
  \begin{pmatrix}
    \phantom{-}\sin(x) cos(y) \\
- \cos(x) sin(y) 
    \end{pmatrix} \\
  \pressure(\bm{x}, t) &= \exp(-4 \mu t) \, \frac{1}{4} \left( \cos(2x) + \cos(2y) \right) + C
  \end{split}
\end{align}
\todo{Constants!}
$C = 100/\gamma$ \cite{dumbser2016high}

\section{Atmospheric Scenarios}
The scenario is described in terms of potential temperature $\potT$
\begin{equation}
  \potT = T \frac{p_0}{p}^{R/c_p},
\end{equation}
where $p_o = \SI{10e5}{\Pa}$ is the reference pressure.
Solving for $T$ leads to
\begin{equation}
  \label{eq:potTToT}
  T = \potT (\frac{p}{p_0})^{R/c_p}.
\end{equation}
To allow for an easier description of the initial conditions we split the $\potT$ into a background state $\backgroundPotT$ and a pertubation $\pertubationPotT$
\begin{equation}
  \label{eq:potT-split}
  \potT = \backgroundPotT + \pertubationPotT.
\end{equation}

The initial conditions for pressure and density can be computed directly from the \textsc{pde} once few assumptions have been given.
We assume that $\backgroundPotT$ is constant over the entire domain.
Inserting the definition of potential temperature, given by \cref{eq:potTToT}, into the equation of state (see \cref{eq:eos}) leads us to
\begin{equation}
  \pressure(z) = \Qrho(z) R {\left( \backgroundPotT \frac{p}{p_0}\right)}^{R/c_p},
\end{equation}
which we solve for $p(z)$.
After some algebra and simplifying, we arrive at
\begin{equation}
 p(z) = p_0 \left( \frac{R \backgroundPotT \rho(z)}{p_0} \right)^{\gamma},
\end{equation}
or equivalently
\begin{equation}
  \label{eq:eos-potT}
\Qrho(z) = \frac{p_{0}^{\frac{R}{c_{p}}} p^{\frac{1}{\gamma}}{\left (z \right )}}{R \backgroundPotT}.
\end{equation}
We want to consider flow in hydrostatic equilibrium, i.e.\ flow where the force originating from the pressure-gradient is exactly balanced by gravitation.
Inserting this assumption in the momentum equation in $z$-direction and using \cref{eq:eos} once again, we arrive at the ordinary differential equation (\textsc{ode})
\begin{align}
  \label{eq:hydrostatic-balance-potT}
  \begin{split}
  \frac{d}{d z} p{\left (z \right )} &= -g \Qrho(z)
                                     = 
        - \frac{g p_{0}^{\frac{R}{c_{p}}} p^{\frac{1}{\gamma}}{\left (z \right )}}{R \backgroundPotT}\\
  p{\left (0 \right )} &= p_0
  \end{split}
\end{align}
where we use the reference pressure $p_0$ as the pressure on ground level.
This \textsc{ode} can be now simply solved by separation of variables.
After simplifying, we arrive at the initial condition for pressure
\begin{align}
  \label{eq:hydrostatic-background-pressure}
p(z) &= \left(\left(1 - \frac{1}{\gamma}\right) \left(C - \frac{g p_{0}^{\frac{R}{c_{p}}} z}{R \backgroundPotT}\right)\right)^{\frac{c_{p}}{R}},\\
\shortintertext{with constant of integration}
 C &= \frac{c_{p} p_{0}^{\frac{R}{c_{p}}}}{R}.
\end{align}

We now compute the pertubated potential temperature $\potT$ and use \cref{eq:potTToT} to convert it to temperature.
We then evaluate the density $\Qrho(\bm{x})$ with \cref{eq:temperature}.
The energy can be computed by inserting the previously computed pressure into the equation of state~(\ref{eq:eos}).

This leaves us with one problem:
The adiabatic no-slip boundary conditions defined by~\vref{eq:no-slip} are no longer valid.
Concretely, we need to impose a viscous heat flux to ensure that the atmosphere stays in hydrostatic balance~\cite{giraldo2008study}
\todo{Flux correct? Fix notation}
\begin{equation}
  \label{eq:atmosphere-bc}
  F^\text{visc} = \kappa \pdv{\overline{T}}{z} =
\frac{R \overline{\theta} \left(\frac{\overline{p}{\left (z \right )}}{p_{0}}\right)^{\frac{R}{c_{p}}} \frac{d}{d z} \overline{p}{\left (z \right )}}{c_{p} \overline{p}{\left (z \right )}}.
\end{equation}
The background pressure can be reconstructed by \cref{eq:hydrostatic-background-pressure}, its derivative is given by \cref{eq:hydrostatic-balance-potT}.

\subsection{Robert's smooth bubble}
\cite{robert1993bubble}
\subsection{Density current}
\cite{straka1993numerical}
%%% Local Variables:
%%% mode: latex
%%% TeX-master: "../main"
%%% End:

>>>>

\chapter{Implementation}\label{chap:implementation}
\todo{Write this chapter!}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: "../main"
%%% End:

### 2018-12-08 writing/thesis/chapters/methods.tex
Some more work.

% Stuff for numerics
\newcommand{\domain}{\Omega}
\newcommand{\broken}{\domain}
\newcommand{\cell}[1][i]{C_{#1}}
\newcommand{\boundary}{\partial \domain}
\newcommand{\sbasis}[1]{\Phi_{#1}}
\newcommand{\testfunction}[1]{\Phi_{#1}}
\newcommand{\normal}{\bm{n}}

@@

\chapter{An ADER-DG scheme for the Navier-Stokes Equations}
\section{Conservative Form}
Standard (hyperbolic) form of conservation law
\begin{equation}
  \label{eq:conservation-law}
 \frac{\partial}{\partial_t}  \Q + \div \flux(\Q) = \source(\bm{x}, t, \Q)
\end{equation}
extend to:
\begin{equation}
  \label{eq:conservation-law-gradient}
 \frac{\partial}{\partial_t}  \Q + \div \flux(\Q, \gradQ) = \source(\bm{x}, t, \Q)
\end{equation}

@@

Inside each cell $\cell$ we represent the solution in terms of the basis function 
\begin{equation}
  \label{eq:cell-approx}
  u(\bm{x}, t^n)_{|\cell[i]} = \hat{u}^n_{i,l} \sbasis{l}{\bm{x}},
\end{equation}
where $l$ is a multi-index, containing one index per spatial dimension.
\todo{dofs $\hat{u}$, local solution, etc.}
For example, $(l = (l_1, l_2))$ for the two dimensional case.
This polynomial is interpolating, i.e.\ \ldots\todo{interpolating?}.
This choice of basis (called \textit{nodal} basis) allows us to easily compute integrals over cells using Gaussian quadrature.
In detail, we use a the Lagrange interpolation polynomials.

@@

\sidetitle{Corrector}
First, we multiply the system \cref{eq:conservation-law} by a test function $\testfunction{i}$ and integrate over the space-time volume $(\cell \times [t^n, t^{n+1}])$.
We arrive at the so called weak formulation of the \textsc{pde} 
\begin{equation}
  \label{eq:weak-pde}
\intdt{\intdcell{
\testfunction{k} \pdv{\bm{Q}}{t}
}}
+
\intdt{\intdcell{
    \testfunction{k} \left( \div{F(Q, \gradient{Q})} \right)
}}
=
\intdt{\intdcell{
    \testfunction{k} S(Q, \bm{x}, t)
}}
\end{equation}
\todo{Include gradient everywhere!}
\todo{Double check the following description}
\newcommand{\stpredictor}[1]{\bm{q}_{#1}}
We now replace the solution $\bm{Q}$ with so called spacetime-predictor $\stpredictor{h}(\bm{x},t)$ into the weak form and write it as a polynomial using the representation of \cref{eq:cell-approx}.
Integrate first by parts in time (note basis here not defined over time), flux divergence by parts in space.
\begin{align}
\begin{split}
\label{eq:corrector}
\left(
\intdcell{
  \testfunction{k} \sbasis{l}
}
\right)
(\bm{u^{n+1} - u^{n}})
+
\left(\intdt{\intdcell{
      \testfunction{k} \Riemann(\bm{q}^-, \bm{q}^+) \cdot \normal
}}\right)
&-\\
\left(\intdt{\intdcell{
    \gradient{\testfunction{k}} \cdot  F(\bm{q})
}}\right)
&=
\left(\intdt{\intdcell{
      \testfunction{k} \source(\bm{q})
}}\right)
\end{split}
\end{align}
The first term motivates an orthogonal basis as this leads to a diagonal mass matrix.

\sidetitle{Predictor}
To derive the predictor, we again take an approximation of our solution in a nodal basis, but now consider polynomials that are defined both in space and time.
\begin{equation}
  \ldots
\end{equation}
We know multiply the conservation law again by a test function (of the same function space as the basis) and arrive at the weak formulation
\begin{equation}
  \label{eq:weak-pde-space-time}
 \ldots
\end{equation}
Similar to the derivation of the corrector, we again integrate the first term by parts in time and the flux divergence in space.
This time we do not use the Riemann solver for the flux boundary term but rather use the discrete solution at time $t$.
Note that this neglects the interaction with neighbouring cells; this is corrected in the corrector step.

\begin{equation}
  \label{eq:space-time-predictor}
  \ldots
\end{equation}
Inserting \cref{eq:cell-approx-space-time} results in a local systems of equations that can be solved in a fixed point iteration scheme.
For details and proof of convergence for the linear case, see (todo).
Write about initial guess (we use naive one anyway, right?)

@@

\section{Compressible Navier Stokes}
Fluid motion can be described by the compressible Navier Stokes equations.
We follow the description in~\cite{dumbser2010arbitrary} for the Navier Stokes part.
The coupling with the advection-diffusion-reaction equations follows~\cite{hidalgo2011ader}.

The vector of conserved quantities is given by
\begin{equation}
  \label{eq:conserved-variables}
 \Q = \left( \Qrho, \Qj, \QE \QZ \right).
\end{equation}
where $\Qrho$ is the density, $\Qj$ is the momentum, $\QE$ the energy density and $\QZ$ is the mass fraction of the chemical reactant.

@@

The hyperbolic flux is given by
\begin{equation}
  \label{eq:hyper-flux}
  \hyperFlux(\Q) = 
  \begin{pmatrix}
    \Qj \\
    \Qv  \otimes \Qj + \bm{I} \pressure  \\
    \Qv \cdot (\bm{I} \QE + \bm{I} \pressure) \\
    \Qj \QZZ
  \end{pmatrix},
\end{equation}
which are the Euler equations coupled with an advection equation.
\newcommand{\diffCoeff}{\varepsilon}
The viscous flux is
\begin{equation}
  \label{eq:visc-flux}
  \viscFlux(\Q, \gradQ) = 
  \begin{pmatrix}
    -\diffCoeff \gradient{\Qrho}\\
    \stressT (\Q, \gradQ)  \\
    \Qv \cdot \stressT (\Q, \gradQ) - \kappa \gradient{T}\\
    -\diffCoeff \gradient{\QZ}
  \end{pmatrix},
\end{equation}

% \begin{equation}
%   F(Q, \nabla Q) = 
%   \begin{pmatrix}
%     \Qj \\
%     \Qv  \otimes \Qj + \bm{I} \pressure + \stressT (Q, \nabla Q)  \\
%     \Qv \cdot (\bm{I} \QE + \bm{I} \pressure + \stressT (Q, \nabla Q)) - \kappa \nabla T
%   \end{pmatrix},
% \end{equation}
where $\pressure$, $\stressT$ and $(\kappa \nabla T)$ denote the pressure, stress tensor and heat flux respectively.
Temperature is denoted by $T$.

We close the system with the equation of state of an ideal gas
\begin{equation}
  \label{eq:eos}
  \pressure = (\gamma - 1) \left(\QE - \frac{1}{2} \left(\Qv \cdot \Qj \right)  - q_0 \QZ \right).
\end{equation}
%The term $\q_0 \QZ$ corresponds to the chemical energy.
The temperature $T$ relates to pressure and density by the ideal gas law
\begin{equation}
  \label{eq:temperature}
 \frac{\pressure}{\Qrho} = RT,
\end{equation}
where $R$ is the specific gas constant.

We further introduce the ratio of specific heats $\gamma$ and the heat fraction for constant volume $c_v$ and constant volume $c_r$.
These constants are all fluid dependent and relate to each other by
\begin{align}
  \begin{split}
  c_v &= \frac{1}{\gamma - 1} R \\
  c_p &= \frac{\gamma}{\gamma - 1} R\\
  R &= c_p - c_v\\
  \gamma &= \frac{c_p}{c_v}.
  \end{split}
\end{align}
\todo{Cite sth!}

We can finally define the heat conduction coefficient $\kappa$
\begin{equation}
  \label{eq:heat-conduction-coeff}
  \kappa = \frac{\mu \gamma}{\Pr} \frac{1}{\gamma - 1} R
\end{equation}
where Prandtl number $\text{Pr}$ again depends on the fluid.

@@

The viscous effects are modeled by the stress tensor
\begin{equation}
  \label{eq:stress-tensor}
  \stressT(Q, \nabla Q) = \left(\nicefrac{2}{3} \mu \nabla \cdot \Qv \right) -
  \mu \left( \nabla (\Qv) + \nabla ( \Qv )^\intercal \right).
\end{equation}

@@

\section{Adaptive Mesh Refinement}
\label{sec:amr}
We use adaptive mesh refinement (\textsc{amr}).

Use total variation (\textsc{tv}) given by
\begin{equation}
  \label{eq:tv}
  \operatorname{TV}\left( f \right) =
\intdcell{ \vert \gradient{f} \vert }
\end{equation}

@@

Criterion with threshold.
Justify by Chebychev's inequality?

>>>>

% Stuff for numerics
\newcommand{\domain}{\Omega}
\newcommand{\broken}{\domain}
\newcommand{\cell}[1][i]{C_{#1}}
\newcommand{\boundary}{\partial \domain}
\newcommand{\sbasis}[1]{\Phi_{#1}}
\newcommand{\stbasis}[1]{\Phi_{#1}}
\newcommand{\testfunction}[1]{\Phi_{#1}}
\newcommand{\normal}{\bm{n}}

@@

\chapter{An ADER-DG scheme for the Navier-Stokes Equations}\label{chap:methods}
\section{Conservative Form}
Standard (hyperbolic) form of conservation law
\begin{equation}
  \label{eq:conservation-law}
 \frac{\partial}{\partial_t}  \Q + \div \flux(\Q) = \source(\bm{x}, t, \Q)
\end{equation}
extend to:
\begin{equation}
  \label{eq:conservation-law-gradient}
 \frac{\partial}{\partial_t}  \Q + \div \flux(\Q, \gradQ) = \source(\bm{x}, t, \Q)
\end{equation}

@@

Inside each cell $\cell$ we represent the solution in terms of the basis function 
\begin{equation}
  \label{eq:cell-approx}
  u(\bm{x}, t^n)_{|\cell[i]} = \hat{u}^n_{i,l} \sbasis{l}(\bm{x}),
\end{equation}
where $l$ is a multi-index, containing one index per spatial dimension.
\todo{dofs $\hat{u}$, local solution, etc.}
For example, $(l = (l_1, l_2))$ for the two dimensional case.
This polynomial is interpolating, i.e.\ \ldots\todo{interpolating?}.
This choice of basis (called \textit{nodal} basis) allows us to easily compute integrals over cells using Gaussian quadrature.
In detail, we use a the Lagrange interpolation polynomials.

@@

\newcommand{\dsol}{\bm{u}}
\newcommand{\stpredictor}[1][h]{\bm{q}_{#1}}
\sidetitle{Corrector}
First, we multiply the system \cref{eq:conservation-law} by a test function $\testfunction{i}$ and integrate over the space-time volume $(\cell \times [t^n, t^{n+1}])$.
We arrive at the so called weak formulation of the \textsc{pde} 
\begin{equation}
  \label{eq:weak-pde}
\intdt{\intdcell{
\testfunction{k} \pdv{\Q}{t}
}}
+
\intdt{\intdcell{
    \testfunction{k} \left( \div{F(\Q, \gradQ} \right)
}}
=
\intdt{\intdcell{
    \testfunction{k} S(\Q, \bm{x}, t)
}}
\end{equation}
\todo{Include gradient everywhere!}
\todo{Double check the following description}
We now replace the solution $\bm{Q}$ with so called spacetime-predictor $\stpredictor (\bm{x},t)$ into the weak form and write it as a polynomial using the representation of \cref{eq:cell-approx}.
Integrate first by parts in time (note basis here not defined over time), flux divergence by parts in space.
\begin{align}
\begin{split}
\label{eq:corrector}
\left(
\intdcell{
  \testfunction{k} \sbasis{l}
}
\right)
(\bm{u^{n+1} - u^{n}})
&+
\left(\intdt{\intdcell{
      \testfunction{k} \Riemann(\stpredictor^-, \stpredictor^+) \cdot \normal
}}\right)
-\\
\left(\intdt{\intdcell{
    \gradient{\testfunction{k}} \cdot  \flux(\stpredictor)
}}\right)
&=
\left(\intdt{\intdcell{
      \testfunction{k} \source(\stpredictor)
}}\right)
\end{split}
\end{align}
The first term motivates an orthogonal basis as this leads to a diagonal mass matrix.

\sidetitle{Predictor}
To derive the predictor, we again take an approximation of our solution in a nodal basis, but now consider polynomials that are defined both in space and time
\begin{equation}
  \label{eq:cell-approx-time}
  q(\bm{x}, t^n)_{|\cell[i]} = \hat{u}^n_{i,l} \stbasis{l}(\bm{x}, t).
\end{equation}
We now multiply the conservation law again by a test function (of the same function space as the basis) and arrive at the weak formulation
\begin{equation}\label{eq:weak-pde-space-time}
\intdt{\intdcell{
    \testfunction{k}(\bm{x}, t)
    \pdv{\stpredictor}{t}
}}
+
\intdt{\intdcell{
    \testfunction{k}{\bm{x}, t}
    \left(
      \divergence{\flux(\stpredictor)}
    \right)
}}
=
\intdt{\intdcell{
  \testfunction{k} \source(\stpredictor)
}}.
\end{equation}
Similar to the derivation of the corrector, we again integrate the first term by parts in time and the flux divergence in space.
This time we do not use the Riemann solver for the flux boundary term but rather use the discrete solution at time $t$.
Note that this neglects the interaction with neighbouring cells; this is corrected in the corrector step.
\todo{Replace $\bm{x}$ with a macro some other letter. Is also coordinate!}

\begin{equation}
  \label{eq:space-time-predictor}
0
  + \ldots
\end{equation}
Inserting \cref{eq:cell-approx-space-time} results in a local systems of equations that can be solved in a fixed point iteration scheme.
For details and proof of convergence for the linear case, see (todo).
Write about initial guess (we use naive one anyway, right?)

@@

\section{Reactive Compressible Navier Stokes Equations}
Fluid motion can be described by the compressible Navier Stokes equations.
We follow the description in~\cite{dumbser2010arbitrary} for the Navier Stokes part.
The coupling with the advection-diffusion-reaction equation follows~\cite{hidalgo2011ader}.

The vector of conserved quantities is given by
\begin{equation}
  \label{eq:conserved-variables}
 \Q = \left( \Qrho, \Qj, \QE, \QZ \right).
\end{equation}
where $\Qrho$ is the density, $\Qj$ is the momentum, $\QE$ the energy density and $\QZ$ is the mass fraction of the chemical reactant.

@@

\newcommand{\diffCoeff}{\varepsilon}
\newcommand{\hyperFluxDef}{
  \begin{pmatrix}
    \Qj \\
    \Qv  \otimes \Qj + \bm{I} \pressure  \\
    \Qv \cdot (\bm{I} \QE + \bm{I} \pressure) \\
    \Qj \QZZ
  \end{pmatrix}
}

\newcommand{\viscFluxDef}{
  \begin{pmatrix}
     -\diffCoeff \gradient{\Qrho}\\
     \stressT (\Q, \gradQ)  \\
     \Qv \cdot \stressT (\Q, \gradQ) - \kappa \gradient{T}\\
     -\diffCoeff \gradient{\QZ}
   \end{pmatrix}
}

The hyperbolic flux is given by
\begin{equation}
  \label{eq:hyper-flux}
  \hyperFlux(\Q) = \hyperFluxDef,
\end{equation}
which are the Euler equations coupled with an advection equation.
The viscous flux is
\begin{equation}
  \label{eq:visc-flux}
  \viscFlux(\Q, \gradQ) = \viscFluxDef.
\end{equation}

Complete \textsc{pde}:
\begin{equation}
  \pdv{}{t}
  \underbrace{
  \begin{pmatrix}
    \Qrho\\
    \Qj\\
    \QE\\
    \QZ
    \end{pmatrix}}_{\Q}
  + 
  \divergence{
  \left(
   \underbrace{\hyperFluxDef}_{\hyperFlux(\Q)}
+
\underbrace{\viscFluxDef}_{\viscFlux(\Q, \gradQ)}
  \right)}
 =
  \underbrace{
  \begin{pmatrix}
    S_{\Qrho\phantom{\Qrho}}\\
    S_{\Qj}\\
    S_{\QE}\\
    S_{\QZ}
    \end{pmatrix}}_{\source(\Q, \bm{x}, t)}
\quad
 \begin{array}{l}
 \text{mass cons.} \\
 \text{momentum cons.} \\
 \text{energy cons.} \\
 \text{cont.\ gas} 
\end{array}
\end{equation}

where $\pressure$, $\stressT$ and $(\kappa \nabla T)$ denote the pressure, stress tensor and heat flux respectively.
Temperature is denoted by $T$.

We close the system with the equation of state of an ideal gas
\begin{equation}
  \label{eq:eos}
  \pressure = (\gamma - 1) \left(\QE - \frac{1}{2} \left(\Qv \cdot \Qj \right)  - q_0 \QZ \right).
\end{equation}
The term $q_0 \QZ$ corresponds to the chemical energy, where $q_0$ is the heat release.
The temperature $T$ relates to pressure and density by the ideal gas law
\begin{equation}
  \label{eq:temperature}
 \frac{\pressure}{\Qrho} = RT,
\end{equation}
where $R$ is the specific gas constant.

We further introduce the ratio of specific heats $\gamma$ and the heat fraction for constant volume $c_v$ and constant volume $c_r$.
These constants are all fluid dependent and relate to each other by
\begin{align}
  \label{eq:fluid-constants}
  \begin{split}
  c_v &= \frac{1}{\gamma - 1} R \\
  c_p &= \frac{\gamma}{\gamma - 1} R\\
  R &= c_p - c_v\\
  \gamma &= \frac{c_p}{c_v}.
  \end{split}
\end{align}
\todo{Cite sth!}
We can finally define the heat conduction coefficient $\kappa$
\begin{equation}
  \label{eq:heat-conduction-coeff}
  \kappa = \frac{\mu \gamma}{\Pr} \frac{1}{\gamma - 1} R
\end{equation}
where Prandtl number $\text{Pr}$ again depends on the fluid.

@@

The viscous effects are modeled by the stress tensor
\begin{equation}
  \label{eq:stress-tensor}
  \stressT(Q, \nabla Q) = \left(\frac{2}{3} \mu \nabla \cdot \Qv \right) -
  \mu \left( \nabla (\Qv) + \nabla ( \Qv )^\intercal \right).
\end{equation}

@@

\section{Adaptive Mesh Refinement \textit{\&} Finite Volume Limiting}\label{sec:grid}
Grid, spacefilling curve, and so on.

\subsection{Adaptive Mesh Refinement}\label{sec:amr}
We use adaptive mesh refinement (\textsc{amr}).

We use the total variation (\textsc{tv}) over an arbitrary indicator variable given by
\begin{equation}
  \label{eq:tv}
  \operatorname{TV}\left[ f(\bm{x}] \right] =
\Vert \left(\intdcell{ \vert \gradient{f \left( \bm{x} \right)} \vert } \right) \Vert_1
\end{equation}

@@

Criterion with threshold.
Justify by Chebychev's inequality?
\todo{Cite this!}
This criterion can be justified by Chebychev's inequality, which states that
\begin{equation}
  \label{eq:chebychev}
  \mathbb{P}(\vert X - \mu \vert \geq c \sigma) \leq \frac{1}{c^2}
\end{equation}
for an arbitrary distribution with mean $\mu$ and standard deviation $\sigma$ probability $\mathbb{P}$.

\subsection{Limiting}\label{sec:limiting}
Higher order \textsc{dg} methods cannot cope with discontinuous solutions.
Even worse, in the case of non-linear fluxes, discontinuities can appear even from smooth initial data.
A classical way of dealing with this problem is the usage of so called limiters.
The idea is to smooth out steep gradients and thus eliminate discontinuities.\todo{It doesnt work like this exactly..}

A relatively recent way of dealing with this problem is the finite volume subcell limiter.
Our discussion follows the one in~\cite{dumbser2016simple}.
This limiting is an a posteriori limiting, which means that we first evaluate a timestep with an unlimited \textsc{ader-dg} method and then check, whether our solution is \enquote{correct}.
In our case, we check whether the solution is a valid floating point number.
We also check whether it is physically admissible.
In our case we use the criterion
\begin{equation}
  \label{eq:limiting-physical}
  \operatorname{is-admissible}(\Q) =
  \begin{cases}
    \text{true} & \Qrho > 0 \land \QE > 0 \\
    \text{false} & \text{otherwise}.
  \end{cases}
\end{equation}
\todo{Mention dmp if I end up using it.}

### 2018-12-08 writing/thesis/chapters/results.tex
Some more work.

\chapter{Results}
\section{Convergence Test}
\section{Accurate CFD}
\section{Accurate Clouds}
\section{AMR Efficiency}
\section{Reactive Euler \textit{\&} Navier Stokes}

>>>>

\chapter{Results}\label{chap:results}
\section{Convergence Test}
Before computing the error we first need to define the $L_p$ norms for a $p > 0$ by
% notation: https://en.wikipedia.org/wiki/Lp_space#Lp_spaces
\begin{equation}
  \label{eq:Lp-nrom}
  \Vert f(x) \Vert = \left( \int_K \vert f(x) \vert^p d\mu  \right)^{1/p}.
\end{equation}
We start by a known analytical solution $f(x)$ and compare it to our approximation $\hat{f}(x)$.
Let $Q$ and $\hat{Q}$ denote the analytical and approximate solution at node.
We start by computing the point-wise error, and compute the cell-wise error from this.
Observe that for our broken space $\Omega$
\begin{equation}
  \label{eq:lp-norm-broken}
 \Vert f(x) \Vert = \sum_{K \in \Omega} \Vert f_K(x) \Vert_p.
\end{equation}

We are now ready to integrate the error for each cell.
The output contains the position and conserved variables of each point.
Each cell consists of $(N + 1)^d$ nodes, each associated with a quadrature weight $w_i$.
The quadrature weights are associated with the reference cuboid.
We thus need to map our element to the reference element, before we can compute the integral.

Let $V$ denote the volume (or area) of each cell.
For example, in three dimensions $V = \Delta x \, \Delta y \, \Delta z$.
The final result is then
\begin{equation}
  \Vert f_K(x) \Vert = \left( V \sum_{\bm{i}} \vert f(\bm{x}_{\bm{i}}) \vert^p w_{{\bm{i}}}  \right)^{1/p},
\end{equation}
where we used the mapping to the reference triangle (integration by substitution).

\section{Accurate CFD}
\section{Accurate Clouds}
\section{AMR Efficiency}
\section{Reactive Euler \textit{\&} Navier Stokes}

### 2018-12-08 writing/thesis/chapters/scenarios.tex
Some more work.

\chapter{Scenarios}\label{sec:scenarios}
\section{Numerical Convergence Test}
We first define the $L_p$ norms for $p > 0$ by
% notation: https://en.wikipedia.org/wiki/Lp_space#Lp_spaces
\begin{equation}
  \label{eq:Lp-nrom}
  \Vert f(x) \Vert = \left( \int_K \vert f(x) \vert^p d\mu  \right)^{1/p}.
\end{equation}
We start by a known analytical solution $f(x)$ and compare it to our approximation $\hat{f}(x)$.
Let $Q$ and $\hat{Q}$ denote the analytical and approximate solution at node.
We start by computing the point-wise error, and compute the cell-wise error from this.
Observe that for our broken space $\Omega$
\begin{equation}
  \label{eq:lp-norm-broken}
 \Vert f(x) \Vert = \sum_{K \in \Omega} \Vert f_K(x) \Vert_p.
\end{equation}

We are now ready to integrate the error for each cell.
The output contains the position and conserved variables of each point.
Each cell consists of $(N + 1)^d$ nodes, each associated with a quadrature weight $w_i$.
The quadrature weights are associated with the reference cuboid.
We thus need to map our element to the reference element, before we can compute the integral.

Let $V$ denote the volume (or area) of each cell.
For example, in three dimensions $V = \Delta x \, \Delta y \, \Delta z$.
The final result is then
\begin{equation}
  \Vert f_K(x) \Vert = \left( V \sum_{\bm{i}} \vert f(\bm{x}_{\bm{i}}) \vert^p w_{{\bm{i}}}  \right)^{1/p},
\end{equation}
where we used the mapping to the reference triangle (integration by substitution).

Following~\cite{dumbser2010arbitrary}, the scenario can be described in primitive variables by setting
\begin{align}
  \pressure &= \pressure_0 \cos( \bm{k} \bm{x} - \omega t ) + \pressure_b, \\
  \Qrho &= \Qrho_0 \sin (\bm{k} \bm{x} - \omega t) + \Qrho_b, \\
  \Qv &= \bm{v_0} \sin(\bm{k} \bm{x} - \omega t).
\end{align}
We set the constants to \( \left(  \pressure_0 = 0.1, \pressure_b = \gamma^{-1}, \Qrho_0 = 0.5, \Qrho_b = 1, 
\bm{v_0} = 0.25 {(1,1)}^\intercal, \bm{k} = \pi/5 {(1,1)}^\intercal \right) \).

@@

\section{Classical Scenarios}

\subsection{The first Stokes Problem}
\cite{dumbser2010arbitrary}
\subsection{Taylor-Green Vortex}
\begin{align}
  \begin{split}
  \Qv(\bm{x}, t) &= \exp(-2 \mu t)
  \begin{pmatrix}
    \phantom{-}\sin(x) cos(y) \\
- \cos(x) sin(y) 
    \end{pmatrix} \\
  \pressure(\bm{x}, t) &= \exp(-4 \mu t) \, \frac{1}{4} \left( \cos(2x) + \cos(2y) \right) + C
  \end{split}
\end{align}
\todo{Constants!}
$C = 100/\gamma$ \cite{dumbser2016high}

>>>>

\chapter{Scenarios}\label{sec:scenarios}
Unless otherwise noted, we use the constants
\begin{align}
\begin{split}
  \gamma &= 1.4, \\
  \Pr &= 0.7, \\
  c_v &= 1.0
\end{split}
\end{align}
The other constants follow from the definition of \cref{eq:fluid-constants}.

We describe some scenarios in terms of the primitive variables $(\Qrho, \Qv, \pressure)$, these can be converted to the conservative variables by the definition of momentum and the definition of energy.   
\section{Numerical Convergence Test}
Following~\cite{dumbser2010arbitrary}, the scenario can be described in primitive variables by setting
\begin{align}
\begin{split}
  \pressure(\bm{x}, t) &= \pressure_0 \cos( \bm{k} \bm{x} - \omega t ) + \pressure_b, \\
  \Qrho(\bm{x}, t) &= \Qrho_0 \sin (\bm{k} \bm{x} - \omega t) + \Qrho_b, \\
  \Qv(\bm{x}, t) &= \bm{v_0} \sin(\bm{k} \bm{x} - \omega t).
\end{split}
\end{align}
We set the constants to \( \left(  \pressure_0 = 0.1, \pressure_b = \gamma^{-1}, \Qrho_0 = 0.5, \Qrho_b = 1, 
\bm{v_0} = 0.25 {(1,1)}^\intercal, \bm{k} = \pi/5 {(1,1)}^\intercal \right) \).

@@

\section{Classical Scenarios}
In addition to the convergence test, we use standard \textsc{cfd} scenarios to evaluate our simulations.

\subsection{The first Stokes Problem}
\cite{dumbser2010arbitrary}
\subsection{Taylor-Green Vortex}
\begin{align}
  \label{eq:taylor-green}
  \begin{split}
  \Qrho(\bm{x}, t) &= 1\\
  \Qv(\bm{x}, t) &= \exp(-2 \mu t)
  \begin{pmatrix}
    \phantom{-}\sin(x) cos(y) \\
- \cos(x) sin(y) 
    \end{pmatrix} \\
  \pressure(\bm{x}, t) &= \exp(-4 \mu t) \, \frac{1}{4} \left( \cos(2x) + \cos(2y) \right) + C
  \end{split}
\end{align}
\todo{Constants!}
$C = 100/\gamma$ \cite{dumbser2016high}

As initial conditions we use \cref{eq:taylor-green} at time $(t = 0)$ and as boundary conditions we impose the solution and the corresponding gradients.
Note that this solution is not exactly correct, as it assumes an incompressible flow.
Due to the low mach number a compressible flow approaches this solution.
